Modeling How Evolution in One Trait is Affected by Adaptation in Another

### 2017-07-20 manuscript.tex
more updates for paper

\section*{Introduction}
Natural selection frequently acts on multiple traits simultaneously. A single trait’s response to selection leads to a shift in the population mean towards the fittest phenotype.  The response to selection with multiple traits involves both direct and indirect effects that together determine their evolution \cite{Scarcelli23102007,Lovell2013,Wagner2011}.  Indirect effects arise from genetic mechanisms that underlie trait expression which may be associated or shared in a manner that enforces specific patterns of co-expression and give rise to trait correlations.\par

Trait correlations are statistical measures of associated co-expression in traits primarily used in quantitative genetics.  Within the

Evolution with multiple traits can produce much more complex responses in populations as they adapt; a fact that was clear to Darwin (Darwin p. 144-150).  Today, evolutionary biologists are in agreement that these indirect influences stemming from simultaneous trait evolution have important consequences that affect populations on a micro- and macro-evolutionary scale.  Plant and animal breeders have been keen on this fact for much longer, and not surprisingly, our most comprehensive theory of adaptation in complex organisms with multiple traits was developed from the quantitative tools used in by them.  The extension of quantitative genetics to multivariate phenotypes, formulated by Lande (Lande 1979; 1980), provides the first characterization of selection on multiple traits, formally expressed in the breeder’s equation ($\Delta z ̅=G \beta$). The phenotypes of quantitative traits change ($\Delta z$) in response to a selection gradient ($\beta$) acting on them, given a specific set of trait variances and covariance between them ($G$). Lande’s formulation is largely independent of the principles of heredity.  However, genetic correlations are known to be the produce of either pleiotropy, linkage disequilibria, and trans acting elements.\par

>>>>

\section*{Introduction}
Natural selection frequently acts on multiple traits simultaneously. A single trait’s response to selection leads to a shift in the population mean towards the fittest phenotype.  The response to selection with multiple traits involves both direct and indirect effects that together determine their evolution \cite{Scarcelli23102007,Lovell2013,Wagner2011}.  Indirect effects arise from genetic mechanisms that underlie trait expression which may be associated or shared in a manner that enforces specific patterns of co-expression.  These effects can significantly influence the evolution of populations on a micro- and macro-evolutionary scale.  This was partly evident to Darwin (Darwin p. 144-150) through his observations of animal and plant breeding.  Evolutionary biologists since then have explore related questions experimentally and theoretically, but the most comprehensive framework for understanding  among geneticists and breeders .\par

In quantitative genetics, indirect effects stemming from the evolution in genetically associated traits is measured by genetic correlations, or alternatively, genetic covariances, and they form part of the statistical description of phenotypic responses to selection

The extension of quantitative genetics to multivariate phenotypes, formulated by Lande (Lande 1979; 1980), provides the first characterization of selection on multiple traits, formally expressed in the breeder’s equation ($\Delta z = \beta$). The phenotypes of quantitative traits change ($\Delta z$) in response to a selection gradient ($\beta$) acting on them, given a specific set of trait variances and covariance between them ($G$). Lande’s formulation is largely independent of the principles of heredity.  However, genetic correlations are known to be the produce of either pleiotropy, linkage disequilibria, and trans acting elements.\par
